%!TEX encoding = UTF-8 Unicode

%% Projeto em Engenharia Aeroespacial 2023/24
%% Template para relattórios preliminar e final 
%%
%% Based on the example.tex file provided by Tomás Oliveira e Silva
%%
%% Modified by João Paulo Barraca <jpbarraca@ua.pt>
%% Modified by Filipe Manco <filipe.manco@ua.pt>
%% Modified by Pedro Casau <pcasau@ua.pt>

% Add 'final' parameter to remove "Relatório Intermédio" from cover page
\documentclass[11pt,twoside,a4paper]{report}

\usepackage[engineering,deti]{uaThesisTemplate}

%% Optional packages for utf8 encoding and T1 fonts
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

%% For prettier pages
\usepackage{lmodern}
\usepackage{amsmath}

%% We will use english and portuguese
\usepackage[english,portuguese]{babel}

\usepackage{xspace}% used by \sigla

\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\DeclareSIUnit\molar{M}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       DEFINE COVER FIELDS
%%                  START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%% Edit author's name
%% Author's full name
\Author{NEIL ARMSTRONG\\ MICHAEL COLLINS\\ EDWIN ALDRIN}

%%% Edit title in Portuguese
\TitlePT{Apollo 11 Mission Overview}

%%% OPTIONAL: you can write your title in
%%%  other languages. You can repeat the following
%%%  line as many times as you want, or remove it
%%%  if you only want the portuguese title.
%% Optional Titles in additional languages.
%\TitleEN{TITLE (MAX 70 CHARACTERS)}

%%% OPTIONAL: edit grant support texts where
%%%  appropriate. Maximum two.
%%%
%%% \GrantText{grant1}{grant2}
%% Optional Grant support texts
%\GrantText{Texto Apoio financeiro do POCTI no âmbito do III Quadro Comunitário de Apoio.}
%          {Texto Apoio financeiro da FCT e do FSE no âmbito do III Quadro Comunitário de Apoio.}

%%% OPTIONAL: Add a dedication text.
%% Optional Dedication text
%\Dedication{Dedico este trabalho à minha esposa e filho pelo incansável apoio.}

%%% OPTIONAL: Add an acknowledgement text
%% Optional Acknowledgement text
%\Acknowledgements{Agradeço toda a ajuda a todos os meus colegas e companheiros.}

%%% Edit the keywords and abstract in Portuguese
%%  Keywords and Abstract in Portuguese
\Abstract{Palavras Chave}{texto livro, arquitetura, história, construção, materiais de construção, saber tradicional.}
         {Resumo}{\selectlanguage{portuguese}Um resumo é um pequeno apanhado de um trabalho mais longo (como uma tese, dissertação ou trabalho de pesquisa). O resumo relata de forma concisa os objetivos e resultados da sua pesquisa, para que os leitores saibam exatamente o que se aborda no seu documento.\\Embora a estrutura possa variar um pouco dependendo da sua área de estudo, o seu resumo deve descrever o propósito do seu trabalho, os métodos que você usou e as conclusões a que chegou.\\Uma maneira comum de estruturar um resumo é usar a estrutura IMRaD. Isso significa: \begin{itemize} \item Introdução \item Métodos \item Resultados \item Discussão \end{itemize} Veja mais pormenores aqui:\\\texttt{https://www.scribbr.com/dissertation/abstract/}}

%%% OPTIONAL: you can write the keywords and
%%%  abstract in other languages. You can repeat
%%%  the following lines as many times as you want
%%%  or remove them if you only want portuguese.
%%% Don't remove the \selectlanguage statement or you will end with incorrect hyphenation.
%%% You can modify the language to any other supported by Latex
%% Optional Keywords and Abstract in additional languages.
\Abstract{Keywords}{textbook, architecture, history, construction, construction materials, traditional knowledge.}
         {Abstract}{\selectlanguage{english}An abstract is a short summary of a longer work (such as a thesis, dissertation or research paper).\\The abstract concisely reports the aims and outcomes of your research, so that readers know exactly what your paper is about.\\Although the structure may vary slightly depending on your discipline, your abstract should describe the purpose of your work, the methods you’ve used, and the conclusions you’ve drawn.\\One common way to structure your abstract is to use the IMRaD structure. This stands for: \begin{itemize} \item Introduction \item Methods \item Results \item Discussion \end{itemize} Check for more details here:\\\texttt{https://www.scribbr.com/dissertation/abstract/}}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       DEFINE COVER FIELDS
%%                  END
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

%%% By default we consider portuguese to be the
%%%  main language (the last one in the usepackage
%%%  declaration). If you're writing your document
%%%  in english uncomment the following line.
%%% Set the proper language or you will end with incorrect hyphenation!
\selectlanguage{english}

\MakeCover

% Disable addition of ToC, LoF and LoT to the table of contents
\addtocontents{toc}{\protect\setcounter{tocdepth}{-1}}

\tableofcontents

\listoffigures
\listoftables

\newpage

% Restore the default value (chapter-level)
\addtocontents{toc}{\protect\setcounter{tocdepth}{2}}

\chapter{Introduction}
\section{Background}
Context or background of the project.
\section{Problem Statement}
Clearly state the problem that you will address in your project.
\section{Objectives}
Clearly state the objectives or goals of your project.
\section{Significance of the Study}
Explain the importance of your project.
\section{Scope}
Define the scope of your project through requirements and allocate priorities in accordance with the relevance and uncertainty of each one.
\section{Work Breakdown Structure (WBS)}
Define the WBS, including Work Packages (WPs), tasks, Gantt Chart, indentifying risks and mitigation strategy.
\chapter{Literature Review}
\section{Influence of IMPs on LDH growth}

AC46000 contains a diverse range of intermetallic particles (IMPs), including copper-rich phases such as \ce{Al2Cu} and \ce{Al7Cu2Fe}, iron-rich phases such as \ce{Al5FeSi} and \ce{Al15(Fe,Mn)3Si2}, and eutectic silicon particles. To assess the influence of these IMPs on LDH growth and corrosion behavior, it is important to consider their electrochemical potential relative to the aluminium matrix \cite{Gudic2024,Kasinska2020}.

The copper-rich phases are strongly cathodic relative to the Al matrix, with \ce{Al2Cu} exhibiting a potential difference of approximately +400 mV \cite{Gudic2024}. This drives intense localized micro-galvanic coupling, promoting rapid oxygen reduction and a sharp rise in local pH. In the context of LDH growth, these conditions trigger the preferential and rapid precipitation of thick, "flower-like" LDH islands directly over the cathodic sites \cite{Cao2026}. This localized growth is sustained by the anodic dissolution of the surrounding aluminium matrix, commonly referred to as trenching \cite{Bouali2022}. The trenching process generates voids at the periphery of the intermetallic particles where LDH flakes fail to nucleate \cite{Bouali2022}. These uncoated gaps compromise the corrosion resistance of the substrate, as chloride ions and other aggressive species can readily penetrate the exposed defects and initiate pitting corrosion \cite{Bouali2022}. The resulting inhomogeneous LDH film also reduces the adhesion of subsequently applied polymer coatings or paints \cite{Bouali2022}, and in severe cases extensive trenching can render the intermetallic particles prone to detachment, further undermining the integrity of overlying layers \cite{Bouali2022}.

The iron-rich phases are also strongly cathodic, approximately +300 mV above the aluminium matrix \cite{Gudic2024}, and the galvanic coupling they generate similarly leads to inhomogeneous LDH growth and increased susceptibility to corrosion.

Although eutectic silicon particles are likewise cathodic, with a potential difference of approximately +250 mV relative to the Al matrix, they are electrochemically inert and do not act as nucleation sites for LDH. Since they provide no source of the metal cations (\ce{Al^2+/Al^3+}, \ce{Mg^2+}) required for LDH formation, no coating growth is triggered at these sites \cite{Gudic2024,DelOlmo2025}.

\subsection{Pre-Treatment}

Given the detrimental influence of these IMPs on LDH growth homogeneity and overall coating quality, their removal from the alloy surface prior to conversion coating is essential. To this end, various pre-treatment strategies have been developed, but the most widely adopted approach consists of a multistep procedure comprising degreasing, alkaline etching, and acid pickling, applied sequentially \cite{MartinezViademonte2020}.

Degreasing constitutes the first pre-treatment step, with the primary objective of removing surface contamination, most notably oils and lubricants introduced during the manufacturing process. This step is critical to ensure full surface wettability, thereby guaranteeing that subsequent processing stages act uniformly across the substrate \cite{MartinezViademonte2020}. 

The second step, alkaline etching, is aimed at dissolving the native oxide layer that forms spontaneously on aluminium surfaces, as well as removing surface-exposed intermetallic particles. The cathodic IMPs act as highly active sites for oxygen reduction, generating a localised rise in pH at their immediate vicinity. This accelerates the anodic dissolution of the surrounding aluminium matrix, progressively undermining the mechanical support of the particles \cite{Nelson2001}. As the matrix recedes, scallop-shaped voids form at the particle periphery, eventually causing the IMPs to detach from the surface, a process known as selective dealloying \cite{Nelson2001,Jin2020}. Since many of the constituent alloying elements released during this process are insoluble in alkaline media, they accumulate on the surface as a loosely adherent residue known as smut, the removal of which necessitates a subsequent acid pickling step \cite{MartinezViademonte2020,Nelson2001,Jin2020}.

Nitric acid (\ce{HNO3}) is widely used for desmutting because it naturally passivates the aluminum matrix while acting as a powerful oxidant \cite{Jin2020}. It successfully targets and dissolves the dealloyed metallic smut, including copper, iron, and manganese residues. Despite this, Nitric acid alone is ineffective against silicon-rich intermetallic residues \cite{ASM1994}. To remove these, fluorides (such as hydrofluoric acid or ammonium bifluoride) must be added to the pickling bath \cite{Askin1995,Lee2008}. The fluoride ions dissolve the silica and complex the remaining silicon remnants so they can be washed away \cite{Askin1995,Lee2008}. 

\section{Equivalent Circuit Models for EIS}
Determining an appropriate equivalent circuit model (ECM) for electrochemical impedance spectroscopy (EIS) spectra is among the most important steps in impedance analysis, as it enables the deconvolution of overlapping electrochemical processes and their quantification through physically meaningful parameters. One of the most commonly used circuit is the Randles equivalent circuit composed of a resistance in series with a resistance and a capacitor in parallel, as represented in FIGURE \cite{Das2025}. As labeled in the figure, the first resistance is the solution resistance and as the name suggests it represents the resistance of the bulk electrolyte between the reference electrode and the working electrode\cite{MetrohmEIS3}. The second resistance on the other hand is known as the charge-transfer resistance and is a kinetic parameter associated with the energy barrier that electrons must overcome to cross the electrode/electrolyte interface during a Faradaic reaction. It is inversely proportional to the exchange current density, reflecting the inherent "speed" of the electrochemical reaction. The capacitor is known as the double layer capacitor, associated with the double layer fenomena explained earlier. Its value is a function of the electrode potential, the ionic concentration, and the effective surface area of the metal. 

As previously mentioned a high charge-transfer resistance means there is a large kinetic energy barrier preventing electrons from crossing the interface, which correlates directly to a very slow corrosion rate. Conversly an increase in Cdl indicates that the protective film has failed, exposing a larger "film-free" area of bare metal directly to the corrosive electrolyte. Additionally, as the metal dissolves and localized corrosion (like pitting) occurs, the surface becomes physically rougher and more fractal; this increased effective surface area directly drives the double layer capacitance higher. This can be related to the formula of a planar capacitor:

Determining an appropriate equivalent circuit model (ECM) for electrochemical impedance spectroscopy (EIS) spectra is among the most critical steps in impedance analysis, as it enables the deconvolution of overlapping electrochemical processes and their quantification through physically meaningful parameters. One of the most widely used models is the Randles equivalent circuit, which consists of a solution resistance ($R_{\text{s}}$) in series with a parallel combination of a charge-transfer resistance ($R_{\text{ct}}$) and a double-layer capacitance ($C_{\text{dl}}$), as represented in [REFERENCIAR FIGURE] \cite{Das2025}. The $R_{\text{s}}$, represents the ohmic resistance of the bulk electrolyte between the reference electrode and the working electrode \cite{MetrohmEIS3}. The $R_{\text{ct}}$, is a kinetic parameter associated with the energy barrier that must be overcome for electrons to cross the electrode/electrolyte interface during a Faradaic reaction. It is inversely proportional to the exchange current density, thereby reflecting the intrinsic rate of the electrochemical reaction at equilibrium. The $C_{\text{dl}}$, arises from the charge separation at the electrode/electrolyte interface, the electric double layer phenomenon described in [REFERENCIAR EARLIER DOUBLE LAYER], and its magnitude is governed by the electrode potential, the ionic concentration of the electrolyte, and the effective surface area of the electrode.

\begin{equation}
    C = \frac{\epsilon \epsilon_0 A}{d} \label{eq:capacitance}
\end{equation}

Where $\epsilon$ is the permittivity, $\epsilon_0$ is the vacuum permittivity, $A$ is the area of the plates, and $d$ is the separation distance.

%% DESCREVER CPE 
%% FALAR DOS CIRCUITOS MAIS AVANÇADOS DE RANDELS 
%% FALAR DE INHIBITOR EFFICIENCY 
%% FALAR DAS TIME CONSTANTS

\chapter{Methodology}
\section{Determine ECM for EIS}
Determining an appropriate equivalent circuit model (ECM) for electrochemical impedance spectroscopy (EIS) spectra is among the most important steps in impedance analysis, as it enables the deconvolution of overlapping electrochemical processes and their quantification through physically meaningful parameters, such as charge transfer resistance ($R_{\text{ct}}$), film or oxide layer resistance ($R_{\text{ox}}$), and double-layer capacitance ($C_{\text{dl}}$).

Despite its centrality to the analysis, ECM identification remains a laborious task. Commercial software packages such as ZView and Gamry Echem Analyst rely on nonlinear least-squares (NLLS) fitting routines, most commonly the Levenberg-Marquardt algorithm, which are inherently sensitive to the choice of initial parameter estimates. While heuristic approaches exist to guide these estimates, such as reading characteristic resistances from the diameters of semicircular arcs in the Nyquist representation or extracting time constants from Bode phase plots, the process remains largely manual and iterative, demanding considerable operator experience to avoid convergence to local minima or physically unreasonable solutions.

In response to these limitations, a number of open-source, programmatic alternatives have been developed, predominantly in Python. Among these, \texttt{impedance.py} provides a flexible framework for ECM fitting and validation. More recently, tools such as \texttt{AutoECM} and \texttt{AutoEIS} have incorporated machine learning methodologies into the workflow. AutoECM employs supervised classification models such as Random Forest, XGBoost (XGB), and convolutional neural network (CNN) architectures. These are trained on libraries of previously validated circuit-spectrum pairs to predict candidate ECMs for new datasets. AutoEIS, by contrast, adopts a probabilistic framework grounded in Bayesian inference, enabling uncertainty quantification in both circuit structure selection and parameter estimation. A notable limitation shared by both tools, however, is the apparent discontinuation of active maintenance by their respective developers. As a consequence, dependency incompatibilities with more recent versions of core scientific Python libraries have rendered them largely non-functional in current environments.

Given this landscape, \texttt{impedance.py} emerges as the most viable open-source option. In its standard implementation, it operates analogously to commercial software: the user must supply the impedance dataset, an \textit{a priori} circuit string, and initial parameter guesses. Two optional fitting configurations are available. The first, \texttt{weight\_by\_modulus}, applies the modulus of the complex impedance as a frequency-dependent weighting factor which ensures that data points spanning several decades of impedance magnitude contribute proportionally to the objective function. The second, \texttt{global\_opt}, invokes a basin-hopping algorithm, which combines stochastic perturbation of parameter coordinates with local minimization and a Metropolis-type acceptance criterion to escape local minima. Importantly, \texttt{global\_opt} is incompatible with \texttt{weight\_by\_modulus} and given that EIS datasets span several decades in frequency, commonly from 10~mHz to 100~kHz, and correspondingly large ranges in impedance magnitude, modulus weighting is generally considered essential to prevent low-frequency, high-impedance data from disproportionately dominating the fit residuals. 

To overcome the dependence on user-supplied initial guesses while retaining the precision of local optimization, a hybrid fitting framework was developed building upon impedance.py and the Differential Evolution (DE) algorithm. DE is a stochastic, population-based global optimization metaheuristic well-suited to high-dimensional, non-differentiable, and multimodal objective functions. The algorithm evolves a population of candidate solutions across successive generations through three operators: mutation, which generates trial vectors via linear combinations of population members; crossover, which probabilistically exchanges parameter values between candidate and trial vectors; and selection, which retains whichever vector yields a lower objective function value. Crucially, DE requires no gradient information and treats the cost function as a black box, making it particularly appropriate for the complex, rugged parameter landscapes characteristic of multi-element ECM fitting.

In the proposed workflow, the user supplies only physically motivated bounds for each circuit element rather than precise initial estimates. The DE algorithm then explores this parameter space to identify the order of magnitude and approximate value of each component. To accelerate convergence, given that circuit element values typically span many orders of magnitude, the optimization is performed in logarithmic space, transforming the search domain such that each decade is treated uniformly. This logarithmic reparameterization substantially improves the efficiency of the population-based search and reduces sensitivity to the scale disparity between, for example, charge-transfer resistances on the order of kilohms and double-layer capacitances on the order of microfarads. Once the DE stage yields a sufficiently accurate global estimate, the resulting parameter vector is passed as the initial condition to the Levenberg-Marquardt local optimizer within \texttt{impedance.py}, with \texttt{weight\_by\_modulus} activated, to refine the solution and obtain statistically meaningful parameter estimates together with their associated uncertainties. 

\section{Pre-treatments}
As previously discussed, the presence of intermetallic phases (IMPs) on the surface of the aluminum alloy can negatively affect LDH growth and must therefore be removed prior to coating. Two different pre-treatment procedures were tested: a chemical treatment and an industrial treatment.

The first procedure consisted of cleaning at room temperature for \qty{120}{\second} in a \qty{0.1}{\molar} \ce{NaOH} solution, followed by etching for \qty{480}{\second} in a \qty{0.1}{\molar} \ce{HNO3} solution. Each step was followed by rinsing with deionized water and final drying in air.

The second procedure consisted of cleaning in Metaclean T2001 (alkaline cleaner based on metal silicates with complexing and surface-active agents) at approximately \qty{65}{\degreeCelsius} for \qty{25}{\minute}, followed by rinsing with deionized water. The specimens were then etched in TurcoLiquid Aluminetch N2 (sodium hydroxide and phosphate-based solution) at \qty{65}{\degreeCelsius} for \qty{45}{\second}, rinsed with deionized water, and desmutted in TurcoLiquid Smutgo NC (based on ferric sulfate, nitric acid, sodium hydrogendifluoride, and sulfuric acid) at \qty{30}{\degreeCelsius} for approximately \qty{5}{\minute}, before a final rinse with deionized water and drying in air.

\chapter{Results}
Present the results of your project clearly and concisely, using tables and figures where necessary.
\chapter{Discussion}
Interpret the results and discuss their implications.
Compare your findings with previous research.
Discuss any limitations of your study.
\chapter{Conclusion \& Future Work}
Sum up the main findings and their importance, and discuss how they have the WBS going forward.




\bibliographystyle{plain}
\bibliography{references}


\end{document}
